\documentclass[a4paper,10pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage[colorlinks=true, urlcolor=blue]{hyperref}
\usepackage{parskip}

\titleformat{\section}{\large\bfseries}{}{0em}{}[\titlerule]
\titleformat{\subsection}[runin]{\bfseries}{}{0em}{}[---]

\begin{document}

%-------------------- Personal Information --------------------%
\begin{center}
    {\LARGE \textbf{Hamish William Patten}} \\
    \vspace{0.2cm}
    hamish.patten@aktiia.com \,|\, +41 77 497 51 01 \\
    LinkedIn: \href{https://linkedin.com/in/hamishpatten}{linkedin.com/in/hamishpatten} \\
    GitHub: \href{https://github.com/hamishwp}{github.com/hamishwp} \\
    ORCID: \href{https://orcid.org/0000-0001-5600-2097}{0000-0001-5600-2097}
\end{center}

%-------------------- Profile --------------------%
\section*{Profile}

Statistician, physicist and data scientist with interdisciplinary experience spanning humanitarian sciences, disaster impact modelling, cardiology and nuclear fusion. Strong communicator with expertise in statistical programming, spatio-temporal modelling, and disaster risk. Practical experience in international research collaborations and software development, including with the United Nations and humanitarian sector applied to complex societal challenges.

%-------------------- Work / Research Experience --------------------%
\section*{Professional Experience}

\textbf{Statistician}, Aktiia \hfill \textit{2025–present} 
\begin{itemize}[noitemsep]
  \item Development of time-series Deep Learning models to estimate blood pressure from optical signals embedded in wearable devices
  \item General data science, including survival analysis-based predictive modelling of user-retention and performance analysis of interventions to decrease blood pressure 
  \item Clinical trial principal statistician, working in trial design and analysis across 6 different trials
  \item Communication via technical reports and scientific articles on blood pressure-related topics.
\end{itemize}


\textbf{Senior Disaster Analyst}, United Nations Food and Agriculture Organization (FAO) \hfill \textit{2025} 
\begin{itemize}[noitemsep]
  \item Estimate global disaster-related losses in crop yields, harvested areas, and livestock, using non-parametric Bayesian time-series methods for the flagship report \textit{Impact of Disasters on Agriculture and Food Security}.
\end{itemize}


\textbf{Disaster Statistics Expert}, UN Economic \& Social Commission for Asia and the Pacific (ESCAP) \hfill \textit{2024–2025} 
\begin{itemize}[noitemsep]
  \item Investigating the intersection of Civil Registration and Vital Statistics (CRVS) systems and disaster-related statistics, including analysing population census versus register data for population demography and disaster-related direct and indirect excess mortality inference.
\end{itemize}


\textbf{Senior Data Scientist}, Int. Fed. of Red Cross and Red Crescent Societies (IFRC) \hfill \textit{2023–2024} 
\begin{itemize}[noitemsep]
  \item Mapped global hazard events to corresponding impact datasets to uncover patterns in disaster occurrence and outcomes.
  \item Analysed trends to inform policy \& publications, including the \textit{World Disasters Report 2023}.
  \item Supported development of Forecast-based Financing models for climate and disaster resilience.
\end{itemize}


%\textbf{Data Science Teacher (volunteer)}, Powercoders \hfill \textit{2024–present} 
%\begin{itemize}[noitemsep]
%  \item Teaching Python and data science skills to people with refugee or asylum seeker status in Switzerland to support integration and employment.
%\end{itemize}


\textbf{Solutions Architect}, United Nations Office for Disaster Risk Reduction (UNDRR) \hfill \textit{2021–2023} 
\begin{itemize}[noitemsep]
  \item Led a team of 9 (data analysts, DevOps, designers) for the design and development of the \textit{Risk Information eXchange (RiX)} platform.
  \item Provided technical support for research and academic collaborations.
\end{itemize}


\textbf{Postdoctoral Researcher}, Department of Statistics, University of Oxford \hfill \textit{2019–2022} 
\begin{itemize}[noitemsep]
  \item Multi-disciplinary research, covering the topics of epidemiology, ecology and disaster impacts
  \item Specialised in Bayesian models and Monte Carlo methods, including Markov Chain Monte Carlo and Approximate Bayesian Computation - Sequential Monte Carlo
\end{itemize}


\textbf{Data Science Research Fellow}, Internal Displacement Monitoring Centre (IDMC) \hfill \textit{2019–2021}


\textbf{PhD Researcher in Nuclear Fusion}, Ecole Polytechnique Fédérale de Lausanne (EPFL) \hfill \textit{2014–2019}


%-------------------- Education --------------------%
\section*{Education}

\textbf{Ecole Polytechnique Fédérale de Lausanne (EPFL)}, CH: \textit{PhD Physics, 2014 -- 2019}. Thesis: \href{https://www.google.com/url?sa=t&source=web&rct=j&opi=89978449&url=https://infoscience.epfl.ch/bitstreams/5babf910-f6b6-440f-a6b6-18fe337d0765/download&ved=2ahUKEwjkybTBpr6MAxUk0wIHHVGqJe4QFnoECBUQAQ&usg=AOvVaw1UCwZ1WpC_BD8WjKRfwaoY}{\textit{“Development and optimisation of advanced auxiliary ion heating
schemes for 3D fusion plasma devices”}}

\vspace{0.1cm}
\textbf{University of York}, UK: \textit{MSc Fusion Energy, 2013 -- 2014}. Graduated with Distinction; awarded a £6000 scholarship. Top of class; recipient of best thesis prize (mark: 78\%) 

\vspace{0.1cm}
\textbf{Loughborough University}, UK: \textit{BSc Mathematics and Physics, 2008 -- 2012}. First Class Honours (final year grade: 84.5\%). Top of class; awarded £1000 merit-based scholarship

%-------------------- Publications --------------------%
\section*{Publications}

\subsection*{1st \& 2nd Author Scientific Articles}
\begin{itemize}[leftmargin=*,noitemsep]
  \item M. Anderson Loake, H. Patten, et al., `A Bayesian Approach for Earthquake Impact Modelling', \textit{Journal of the Royal Statistical Society Series C: Applied Statistics}, accepted manuscript (2026)
  \item R. Browning, H. Patten, et al., `Bayesian spatiotemporal modelling of political violence and conflict events using discrete-time Hawkes processes', \textit{Journal of the Royal Statistical Society Series A: Statistics in Society}, qnag039 (2026)
  \item H. Patten and Z. Bhaby, `Disasters, the Humanitarian Sector and Statistics', \textit{Annual Review of Statistics and Its Applications}, Invited Review, 13:51-73 (2026)
  \item D. Steinsaltz, H. Patten et al., `Short-term and Mid-term Blood Pressure Variability and Long-term Mortality', \textit{American Journal of Cardiology}, 234:71-78 (2024)
  \item H. Patten et al., `Data-Driven Earthquake Multi-Impact Modelling: A Comparison of Models', \textit{International Journal of Disaster Risk Science}, 15, pp. 421–433 (2024)
  \item Ya. Kolesnichenko, H. Patten et al., \textit{Nuclear Fusion}, 61(4), 046014 (2021)
  \item F. Casson, H. Patten et al., \textit{Nuclear Fusion}, 60, 066029 (2020)
  \item H. Patten et al., `Identification of an optimized heating and fast ion generation scheme for the Wendelstein 7-X stellarator', \textit{Physical Review Letters}, 124, 155001 (2020)
  \item H. Patten et al., \textit{Plasma Physics and Controlled Fusion}, 60, 085009 (2018)
\end{itemize}

\subsection*{Thesis}
\begin{itemize}[leftmargin=*,noitemsep]
  \item H. Patten, `Development and optimisation of advanced auxiliary ion heating
schemes for 3D fusion plasma devices', \textit{PhD Thesis 9167}, EPFL, 2019
\end{itemize}

\subsection*{Reports \& Policy Publications}
\begin{itemize}[leftmargin=*,noitemsep]
  \item Hilo by Aktiia, `A nation under pressure: UK blood pressure report 2025' (2025). Access: \href{https://hilo.com/uk/bp-report-2025/?srsltid=AfmBOoqoZO2R-vYM6bqqg5fRD50CchthXeD-OJmGSbu8x2bMiZ7gvKKW}{Aktiia} 
  \item H. Patten \& N. Richards, `Disaster-related Statistics and Civil Registration and Vital Statistics systems : working paper on the nexus between disaster-related statistics and civil registration and vital statistics (CRVS)', \textit{UN ESCAP - SD/WP/17/March2025}, (2025). Access: \href{https://hdl.handle.net/20.500.12870/7973}{ESCAP}
  \item H. Patten et al., `COP28: The Climate and Environmental Crisis is the IFRC’s Top Global Challenge', \textit{Red Cross Red Crescent Climate Centre}, published 30.11.2023. Access: \href{https://www.climatecentre.org/13011/cop28-the-climate-and-environmental-crisis-is-the-ifrcs-top-global-challenge-by-the-climate-centre/}{IFRC}
  \item \textit{World Disasters Report 2023}, IFRC, published 30.11.2023. Access: \href{https://www.ifrc.org/document/world-disasters-report-2022}{IFRC}
  \item H. Patten, `Rethinking Predictive Analytics for Disaster Resource Allocation', \textit{IDMC GRID Background Paper}, 20.05.2021. Access: \href{https://www.internal-displacement.org/global-report/grid2021/downloads/background_papers/background_paper-analytics.pdf}{IDMC}
  \item \textit{Global Report on Internal Displacement (GRID) 2021}, IDMC, published 20.05.2021. Access: \href{https://api.internal-displacement.org/sites/default/files/publications/documents/grid2021_idmc.pdf}{IDMC}
  \item \textit{Global Report on Internal Displacement (GRID) 2020}, IDMC, published 20.05.2020. Access: \href{https://api.internal-displacement.org/sites/default/files/publications/documents/2020-IDMC-GRID.pdf}{IDMC}
  \item \textit{Women and Girls in Internal Displacement}, IDMC, published 05.03.2020. Access: \href{https://api.internal-displacement.org/sites/default/files/publications/documents/202003-twice-invisible-internally-displaced-women.pdf}{IDMC}
\end{itemize}

%-------------------- Research Interests --------------------%
\section*{Research Interests}

\begin{itemize}[noitemsep]
  \item Bayesian methods for spatio-temporal climate \& disaster risk modelling
  \item Computational social science and the use of statistics in humanitarian contexts
  \item Statistical learning for disaster risk reduction and resilience planning
  \item Alternative data sources: Earth Observation data, population registers, mobile positioning data
  \item Methodological development for Forecast-based Financing and Early Warning Systems
  \item Cross-disciplinary applications of data science in the digital and environmental humanities
\end{itemize}

%-------------------- Skills --------------------%
\section*{Skills}
\textbf{Programming:} R, Python, LaTeX, RMarkdown, RShiny, Leaflet, ArcGIS, Fortran, C++. \\
\textbf{Languages:} fluent in English \& French, B2 in Italian \& Portuguese. \\
\textbf{Other:} High Performance Computing, GIS, SQL and data visualisation.

%-------------------- Conferences and Talks --------------------%
\section*{Conferences and Presentations 2024--2025}
\begin{itemize}[leftmargin=*,noitemsep]
  \item `Dose-dependent associations between systolic blood pressure and user engagement with a wearable cuffless monitor', European Society of Hypertension (ESH), European Meeting on Hypertension and Cardiovascular Protection, Gdansk (2026)
  \item `Hilo Exposure Effect: BP Reduction Interventions', British and Irish Hypertension Conference 2025, Cambridge (2025)
  \item `Applying AI to Classify Hazards from Text', European Geosciences Union (EGU) Conference, April 2025, Austria.
  \item `Disaster \& Conflict Modelling with Bayesian Methods', University of Zurich, Department of Mathematics, January 2025, Switzerland.
  \item `Disaster-related Statistics and Civil Registration and Vital Statistics systems', 4th Global Expert Forum for Producers and Users of Disaster-Related Statistics, United Nations Economic Commission for Africa, October 2024, Ethiopia.
  \item `Geopolitical Analyses of Disaster Impacts: Bias, Modelling and Decision Making', invited session at Royal Statistical Society Conference - September 2024, United Kingdom.
  \item `The use of AI in Disaster Risk', contributed presentation, Humanitarian Networks and Partnerships Weeks (HNPW) May 2024, Switzerland.
  \item `Monty: Combining Global Disaster Impact Databases', invited presentation, Regional EM-DAT Annual Workshop, February 2024, South Africa.
\end{itemize}

%-------------------- Teaching Experience --------------------%
\section*{Teaching Experience}
\begin{itemize}[leftmargin=*,noitemsep]
    \item Teacher: \textbf{Data Science with Python}, PowerCoders \textit{2024-Present} 
    \item Lead Tutor: \textbf{Foundations of Statistical Methods}, University of Oxford \textit{2021-2022} 
    \item Assistant: \textbf{Statistical Programming in R}, University of Oxford \textit{2019-2021} 
    \item Assistant: \textbf{Computational Statistics and Monte Carlo Methods}, University of Oxford \textit{2021-2022} \item Assistant: \textbf{Numerical Physics I \& II}, Ecole Polytechnique Fédérale de Lausanne (EPFL), \textit{2015-2018} 
    \item Assistant: \textbf{Solid State Physics}, York University, \textit{2013-2014} 
    \item Assistant: \textbf{Laboratory Physics}, York University, \textit{2013-2014}     
\end{itemize}
%-------------------- Grants and Awards --------------------%
\section*{Grants and Awards}
\begin{itemize}[leftmargin=*,noitemsep]
  \item `Leveraging the Global Crisis Data Bank to enhance risk analysis and impact forecasts at country level', United Nations Disaster Risk Reduction, 2024, 60,000CHF.
  \item `Meta Unrestricted Research Award', Meta, 2022, \$25,000.
  \item `EPSRC Impact Acceleration Account (IAA)', UK Research and Innovation, 2022, £75,000.
  \item Computational Resources Award, Swiss National Supercomputing Centre, 2018, 250,000 CPU hours.
\end{itemize}

%-------------------- Scientific Committees & Memberships --------------------%
\section*{Scientific Committees \& Memberships}
\begin{itemize}[leftmargin=*,noitemsep]
  \item Fellow of Royal Statistical Society, UK.
  \item Member of the EM-DAT Scientific Committee, University of Louvain, BE.
  \item Member of the Risk Data Library Standard Scientific Committee, World Bank - GFDRR.
  \item Member of the Social Equity and Risk Assessment Working Group, University of Michigan, US.
  \item Member of the Institute of Physics, UK.
\end{itemize}


%-------------------- Extra-curricular --------------------%
\section*{Extra-curricular}
\begin{itemize}[leftmargin=*,noitemsep]
  \item La Soupe Populaire: volunteering weekly with a soup kitchen in Lausanne.
  \item Pottery: specialising in wheel-thrown ceramics.
  \item Long-distance running: competing in marathons, 100kms and infinity trails.
  \item Freediving: training \& competing with `Immersion Profonde' club, Lausanne.
  \item Salsa: following weekly courses from AliaSalsa, Lausanne, since April 2024
\end{itemize}

%-------------------- References --------------------%
\section*{References}

\begin{itemize}[leftmargin=*,noitemsep]
  \item \textbf{Dr. Josep Solà} \\
  Aktiia Sarl, Neuchatêl \\
  \href{mailto:josep@aktiia.com}{josep@aktiia.com}
  \vspace{0.2cm}
  \item \textbf{Prof. David Steinsaltz} \\
  Department of Statistics, University of Oxford \\
  \href{mailto:david.steinsaltz@stats.ox.ac.uk}{david.steinsaltz@stats.ox.ac.uk}
  \vspace{0.2cm}
  \item \textbf{Mr. Justin Ginnetti} \\
  International Federation of Red Cross and Red Crescent Societies (IFRC) \\
  \href{mailto:justin.ginnetti@ifrc.org}{justin.ginnetti@ifrc.org}
\end{itemize}


\end{document}
